comgen-outline

### 2017-09-01 paper/comgen-paper.Rmd
sophies feedback for model section

As adults we readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find me typing.
Generalizations are used to make predictions about properties of instances that an agent has yet to experience [@HumeTHN]. 
From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007].
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$. 
The entities can be objects (a dog), events (a coffee drink), or more fine grained types (me on a given day).
The properties can vary greatly (e.g., having fur, causing jitters, typing). 
Yet the prediction for a new entity can be described as a *predictive probability*, which we will denote throughout the paper as $h$ and refer to as the **prevalence**: 
$h = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$.^[
We denote prevalence as $h$ to emphasize that it is a latent *hypothesis*, and not necessarily an observable frequency.
]

@@

The human capacity to generalize from observations has been studied extensively in cognitive and developmental psychology [@Nisbett1983; @Baldwin1993; @Heit2000].
Observing some number of instances of $k$ with $f$ will update a prior belief distribution into a posterior belief distribution.
As a concrete example, consider the feature \textsc{is friendly} ($f$) for the category \textsc{dog} ($k$): $h = P(x \in \{\text{is friendly}\} \mid x \in \{\text{is a dog}\})$.
If an agent were to encounter some number of individual dogs and note how many of them were friendly ($d$), using Bayes' rule, she could form a posterior belief distribution about the probability that the next dog she encounters will be friendly (in other words, a belief about the friendliness of dogs).

@@

Our ability to use this updated belief $P(h \mid d)$ in order to make predictions about unobserved instances is a computational description of generalization [e.g., @Tenenbaum2006; @Tenenbaum2011].

Generalization from examples shows complex sensitivites to the kind of observations, and even how those observations came to be observed in the first place.
For instance, observing that an indigenous person living on a remote island *is obese* provides considerably less information about the weights of other islanders in comparison to inferences about skin color upon learning that one islander's skin color *is brown* [@Nisbett1983].
This subtlety can be modeled using heirarchical models to provide structure to the prior distribution in a Bayesian model [@Kemp2008].
Observations that are generated by an intentional agent can lead to similarly subtle inferences, like unconfounding otherwise ambiguous causal evidence [@Goodman2009].

Studying generalizations from observational evidence can only take us so far, however, in understanding how the mind understands the world or culture accumulates across generations.
Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *starting at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*).
For abstract, generalizable knowledge to remain in a culture, it must be able to be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations.

@@

If generalization from observations can be described using probability, then it seems  natural to consider the semantic content of an utterance that conveys generalization also to be relating to probability. 
One very simple semantic theory that has been used for describing quantified statements (e.g., "some", "all") is a *threshold semantics*.
For generalizations, the threshold would apply to the subjective probability or *prevalence* $h$:

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the **prevalence prior** $P(h)$ plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual accounts of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Contrary to these accounts, in our model, conceptual structure is represented in listeners' knowledge of the world $P(h)$ and not in the semantic content of the utterance $u$.

@@

In this paper, we are interested in explaining the variable *endorsements* of generalizations (e.g., "Birds lay eggs" is good; "Birds are female" is not) based on some known or believed prevalence $h$ (e.g., only female birds lay eggs).
Following the Rational Speech Act (RSA) framework [@Goodman2016], we describe an **endorsement model** as a Bayesian speaker's decision of whether or not to say the generalization to the listener $L$ (Eq. \ref{eq:L0}).^[A more general version of this model can relax the assumption that the speaker has access to a specific value of prevalence $h$ that he is trying to communicate.
Rather, the speaker may have a belief distribution over $h$ for the category $k$, corresponding to the speaker's beliefs about the prevalence of the property in the category. 
In this situation, we would define the speaker's utility to be the expected value of the informativity, which integrates over her belief distribution:
$S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{h\sim P_{k}} \ln{ \int_{\theta} L(h, \theta \mid u)} \diff \theta )}$
]

@@

The speaker $S$ doesn't know the threshold $\theta$; rather, he integrates over the likely values that the listener will believe. 
Additionaly, following RSA, the speaker model is assumed to be an approximately rational Bayesian agent with degree with rationality governed by parameter $\lambda$, which acts as a soft-max function.

The speaker's decision to say an utterance (e.g., the generalization) is determined by how well it would communicate $h$ to the listener $L$, in comparison to alternative utterances.
To model endorsement, this decision is evaluated in comparison to the alternative of *not saying the generalization*;
with just two alternatives (say the utterance vs. not), the model can be interpreted as a model of felicity judgments [@Degen2014]. 
Formally, we consider the negative alternative as the option of staying quiet, which carries no information content^[
This alternative can be realized in at least two other ways: the speaker could have said the negation of the utterance (i.e., *The generalization is not true.*) or the negative generalization (i.e., *The opposite is true.*, e.g., *Ks do not have F*).
All results reported are similar for these two alternatives, and we use the alternative of the *silent* utterance for simplicity.
]:

@@

The pair of speaker-listener models presented above correspond to the simplest possible communicative models with an underspecified threshold-meaning for a generalization. 
The *simple but underspecified* hypothesis defines a *family* of models, however, with potentially more levels of recursion (e.g., a listener who thinks about a speaker who thinks about a listener).
These are not of primary theoretical interest to this paper, as *any* formal model that explains the very flexible endorsement patterns of generalizations in language will be a substantial advancement.
However, there are interesting theoretical differences to be found within this family of models (e.g., if the inference about the threshold can be thought of as a *pragmatic inference* or just a mere *literal inference*).
We discuss these alternative formulations and provide formal model comparison between these RSA models in Appendix.

@@

The RSA model defined in the previous section predicts that the endorsement of a generalization depends upon the probability of that utterance vs. silence accurately conveying to a naive listener a prevalence $h$ in a referent-category (e.g., the \% of robins that lay eggs).
This probability is given by $L(h \mid u, \theta)$ (Eq. \ref{eq:S1}) and depends upon  the listener's *a priori* beliefs about the prevalence $h$ in question: $P(h)$ (Eq. \ref{eq:L0}), as well as the referent-category prevalence $h$ (Eq. \ref{eq:S1}).
Predictions for endorsement depend upon how well the listener's posterior on $h$  upon hearing a generalization matches an $h$ that the speaker wants to convey in comparison to how well the listener's prior matches (which would result from the speaker staying silent).
For purposes of illustration, we present different hypothetical priors on prevalence $P(h)$, together with listener $L$ posterior distributions on $h$ after hearing utterances corresponding to (i) the underspecified threshold semantics as well as (ii) a comparison model where the threshold is fixed near-0, corresponding intuitively to the quantifier *some* (Figure \ref{fig:simulations}).

@@

How does the model handle the puzzling case of "Robins lay eggs" vs. "Robins are female"?
In each case, we must consider the prevalence prior for the feature:  $P(h_\text{lay eggs})$ and $P(h_\text{are female})$.
For being female, no matter want animal category a person might think of, the feature is prevalent in the exact same proprotion (i.e., 50\% of X are female; Figure\ \@ref(fig:simulations) bottom row).
For laying eggs, it really depends: If a person thinks of a falcon, then 50\% of that category lays eggs; if a person thinks of a bear, then 0\% of them lay eggs.
This dependency is reflected in the corresponding $P(h)$, which is a bimodal distribution with peaks near-0\% and near-50\% (Figure\ \@ref(fig:simulations) row 3).
For both priors, the interpretation upon hearing the generalization is the same (i.e., roughly 50\% of the category has the feature; Figure\ \@ref(fig:simulations) rows and 5, middle red panel).
The difference in endorsement comes from the comparison to the prior: For being female, the listener is very likely to arrive at 50\% under her prior, but for laying egges, this is not the case. 
In other words, "Robins are female" has little very informational content as the listener's posterior is nearly identical with her prior (i.e., it doesn't matter if the speaker says the generalization or stays silent).
"Robins lay eggs", however, has a lot of informational content.

>>>>

Generalizations are used to make predictions about properties of instances that an agent has yet to experience [@HumeTHN].
People readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find me typing.
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$. 
The entities can be objects (a dog), events (a coffee drink), or more fine grained types (me on a given day).
The properties can vary greatly (e.g., having fur, causing jitters, typing). 
Yet the prediction for a new entity can be described as a *predictive probability*, which we will denote throughout the paper as $h$ and refer to as the **prevalence**: 
$h = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$.^[
We denote prevalence as $h$ to emphasize that it is a latent *hypothesis*, and not necessarily an observable frequency.
]

@@

The human capacity to generalize from observations has been studied extensively in cognitive and developmental psychology [@Nisbett1983; @Baldwin1993; @Heit2000].
From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007].
The Bayesian probability calculus provides a computational description of generalization [e.g., @Tenenbaum2006; @Tenenbaum2011]: Observing some number of instances of $k$ with $f$ will update what an observer believes *a priori* (formally: a prior belief distribution) in order to make predictions about unobserved instances, via a posterior belief distribution.
As a concrete example, consider the feature \textsc{barks} ($f$) for the category \textsc{dog} ($k$): $h = P(x \in \{\text{barks}\} \mid x \in \{\text{is a dog}\})$.
If an agent were to encounter some number of individual dogs and note how many of them bark ($d$), using Bayes' rule, she could form a posterior belief distribution about the probability that the next dog she encounters will bark (in other words, a belief about the barkiness of dogs).

@@

Generalization from examples shows complex sensitivites to the kind of observations, and even how those observations came to be observed in the first place.
For instance, observing that an indigenous person living on a remote island *is brown and obese* provides considerably more information about the skin color of other islanders than about their weight [@Nisbett1983].
This subtlety can be modeled by incorporating heiarchical structure into the prior belief distribution of a Bayesian model, allowing the model to represent and reason flexibly about different kinds of properties [@Kemp2008].
Observations that are generated by an intentional agent can lead to similarly subtle inferences, like unconfounding otherwise ambiguous causal evidence [@Goodman2009].

Studying generalizations from observational evidence can only take us so far, however, in understanding how the mind understands the world or culture accumulates across generations.
Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*).
For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations.

@@

If generalization from observations can be described using probability, then it seems  natural to consider the semantic content of an utterance that conveys generalization also to be relating to probability. 
One very simple semantic theory is a *threshold semantics*: an utterance is true is some quantity or *degree* is above a threshold.
For generalizations, the threshold would apply to the subjective probability or *prevalence* $h$:

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the **prevalence prior** $P(h)$ plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual accounts of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Contrary to these accounts, in our model, conceptual structure is represented in listeners' knowledge of the world $P(h)$ and not in the semantic content of the utterance $u_{gen}$.

@@

In this paper, we are interested in explaining the variable *endorsements* of generalizations (e.g., "Birds lay eggs" is good; "Birds are female" is not) based on some known or believed prevalence $h$ (e.g., only female birds lay eggs).
Following the Rational Speech Act (RSA) framework [@Goodman2016], we describe a **speaker model** as a Bayesian decision of whether or not to say the generalization to the listener $L$ (Eq. \ref{eq:L0}).^[A more general version of this model can relax the assumption that the speaker has access to a specific value of prevalence $h$ that he is trying to communicate.
Rather, the speaker may have a belief distribution over $h$ for the category $k$, corresponding to the speaker's beliefs about the prevalence of the property in the category. 
In this situation, we would define the speaker's utility to be the expected value of the informativity, which integrates over her belief distribution:
$S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{h\sim P_{k}} \ln{ \int_{\theta} L(h, \theta \mid u)} \diff \theta )}$
]

@@

The speaker $S$ doesn't know the threshold $\theta$; rather, he integrates over the likely values that the listener will believe. 
Additionaly, following RSA, the speaker model is assumed to be an approximately rational Bayesian agent with rationality governed by parameter $\lambda$, which acts as a soft-max function.

The speaker's decision to say an utterance (e.g., the generalization) is determined by how well it would communicate $h$ to the listener $L$, in comparison to alternative utterances.
We specify a single alternative utterance corresponding to *not saying the generalization*:
With just these two alternatives (say the generalization vs. not), the model can be interpreted as a model of felicity judgments or sentence *endorsement* [@Degen2014]. 
Formally, we consider the negative alternative as the option of staying quiet, which carries no information content^[
This alternative can be realized in at least two other ways: the speaker could have said the negation of the utterance (i.e., *The generalization is not true.*) or the negative generalization (i.e., *The opposite is true.*, e.g., *Ks do not have F*).
All results reported are similar for these two alternatives, and we use the alternative of the *silent* utterance for simplicity.
]:

@@

The pair of speaker-listener models presented above correspond to the simplest possible communicative models with an underspecified threshold-meaning for a generalization. 
The *simple but underspecified* hypothesis defines a *family* of models, however, with potentially more levels of recursion (e.g., a listener who thinks about a speaker who thinks about a listener).
These are not of primary theoretical interest to this paper, as *any* formal model that explains the flexible endorsement patterns of generalizations in language will be a substantial advancement.
However, there are interesting theoretical differences to be found within this family of models (e.g., if the inference about the threshold can be thought of as a *pragmatic inference* or just a mere *literal inference*).
We discuss these alternative formulations and provide formal model comparison between these RSA models in Appendix.

@@

The RSA model defined in the previous section predicts that the endorsement of a generalization depends upon the probability that utterance vs. silence accurate conveys to a naive listener a prevalence $h$ in a referent-category (e.g., the \% of robins that lay eggs).
This probability is given by $L(h \mid u, \theta)$ (Eq. \ref{eq:S1}) and depends upon  the listener's *a priori* beliefs about the prevalence $h$ in question: $P(h)$ (Eq. \ref{eq:L0}), as well as the referent-category prevalence $h$ (Eq. \ref{eq:S1}).
Predictions for endorsement depend upon how well the listener's posterior on $h$  upon hearing a generalization matches an $h$ that the speaker wants to convey in comparison to how well the listener's prior matches (which would result from the speaker staying silent).
For purposes of illustration, we present different hypothetical priors on prevalence $P(h)$, together with listener $L$ posterior distributions on $h$ after hearing utterances corresponding to (i) the underspecified threshold semantics as well as (ii) a comparison model where the threshold is fixed near-0, corresponding intuitively to the quantifier *some* (Figure \ref{fig:simulations}).

@@

How does the model handle the puzzling case of "Robins lay eggs" vs. "Robins are female"?
In each case, we must consider the prevalence prior for the feature:  $P(h_\text{lay eggs})$ and $P(h_\text{are female})$.
For being female, no matter what animal category a person might think of, the feature is prevalent in the exact same proprotion (i.e., 50\% of X are female; Figure\ \@ref(fig:simulations) bottom row).
For laying eggs, it really depends: If a person thinks of a falcon, then 50\% of that category lays eggs; if a person thinks of a bear, then 0\% of them lay eggs.
This dependency is reflected in the corresponding $P(h)$, which is a bimodal distribution with peaks near-0\% and near-50\% (Figure\ \@ref(fig:simulations) row 3).
For both priors, the interpretation upon hearing the generalization is the same (i.e., roughly 50\% of the category has the feature; Figure\ \@ref(fig:simulations) rows and 5, middle red panel).
The difference in endorsement comes from the comparison to the prior: For being female, the listener is very likely to arrive at 50\% under her prior, but for laying egges, this is not the case. 
In other words, "Robins are female" has little very informational content as the listener's posterior is nearly identical with her prior (i.e., it doesn't matter if the speaker says the generalization or stays silent).
"Robins lay eggs", however, has a lot of informational content.
